Modal type theory based on the intuitionistic epistemic logic

### 2018-02-15 article.tex
added acknowledgment
added some beauty again

\begin{itemize}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$;
  \item If $\Gamma \vdash M : A$ and $\Delta$ is a permutation of $\Gamma$, then $\Delta \vdash M : A$.
\end{itemize}
\end{lemma}

@@

\begin{ex}
  $ $

  $\phi((\lambda_i x. {\bf let \: pure \:} y = x {\: \bf in \:} x) N) = {\bf let \: pure \:} y = \phi(N) {\: \bf in \:} \phi(N)$
\end{ex}

@@

  $\begin{array}{lll}
  & \phi({\bf pure \:} (M [x := N])) = & \text{By the definition of $\phi$}\\
  & {\bf pure \:} (\phi(M [x := N])) = & \text{Induction hypothesis} \\
  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = & \text{Substitution definition}\\
  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
  \end{array}$

@@

  $\begin{array}{lll}
  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = & \text{Substitution definition}\\
  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = & \text{By the definition of $\phi$}\\
  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = & \text{Induction hypothesis} \\
  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =& \text{Substitution definition} \\
  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
  \end{array}$

@@

Strong endofunctor on a monoidal category is with a natural transformation so called tensorial strength

\begin{equation}
  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
\end{equation}

@@

\xymatrix{
&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{\rho_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(\rho_A)}\\
&&&&& \mathcal{K}A
}
\end{defin}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = f : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda(f) : [\![\Gamma]\!]
\rightarrow[\![B]\!]^{[\![A]\!]}$}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A ]\!] = f : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
\AxiomC{$[\![\Gamma \vdash N : B ]\!] = g : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
\BinaryInfC{$[\![\Gamma \vdash (M, N) : A \times B]\!] = \langle f, g \rangle : [\![\Gamma]\!] \rightarrow
[\![A]\!] \times [\![B]\!]$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash p : A_1 \times A_2]\!] = f : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
[\![A_2]\!]$}
\RightLabel{$i \in \{1,2\}$}
\UnaryInfC{$[\![\Gamma \vdash \pi_i p : A_i]\!] = [\![\Gamma]\!] \xrightarrow{f} [\![A_1]\!] \times
[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
\end{prooftree}

@@

$\begin{array}{lll}
&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \text{By interpretation} \\
&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \text{By the definition of an applicative functor} \\
&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \text{By naturality of $\eta$} \\
&\eta_{[\![A]\!]} \circ [\![M]\!] = & \text{By interpretation} \\
&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
\end{array}$
\end{proof}

@@

$\begin{array}{lll}
&\eta_{\hat{B}} \circ f = & \text{By the definition} \\
&[y, {\bf pure \:} y] \circ [x, M] = & \text{By the definition of composition}\\
&[x, {\bf pure \:} y [y := M]] = & \text{By substitution}\\
&[x, {\bf pure \:} M]& \\
& &\\
&\text{On the other hand:}& \\
&\text{fmap } f \circ \eta_{\hat{A}} = & \text{By the definiton} \\
&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = & \text{By the definition of composition}\\
&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \text{By substitution} \\
&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \text{$\beta$-reduction rule}\\
&[x, {\bf pure \:} M [x := x]] = & \text{By substitution}\\
&[x, {\bf pure \:} M]&
\end{array}$

@@

$\begin{array}{lll}
&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By definition} \\
&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \text{By composition} \\
&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \text{By substitution} \\
&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \text{By $\beta$-reduction rules} \\
&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
\end{array}$

@@

  \begin{itemize}
    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
    \item ${\bf K}(\rho_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = \rho_{{\bf K}\hat{A}}$.
  \end{itemize}
  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
  $\rho = \pi_2$.
\end{lemma}

@@

>>>>

\begin{itemize}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
\end{itemize}
\end{lemma}

@@

@@

  $\begin{array}{lll}
  & \phi({\bf pure \:} (M [x := N])) = & \\
  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$} & \\
  & {\bf pure \:} (\phi(M [x := N])) = & \\
  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
  & {\bf pure \:} (\phi(M) [x := \phi(N)]) = &\\
  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
  & ({\bf pure \:} \phi (M)) [x := \phi(N)]&
  \end{array}$

@@

  $\begin{array}{lll}
  &\phi(({\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \:} N) [y := P]) = &\\
  &\quad\quad\quad\quad\quad\quad \text{Substitution definition}& \\
  &\phi({\bf let \: pure \:} \vec{x} = (\vec{M}[y := P]) {\: \bf in \:} N) = &\\
  &\quad\quad\quad\quad\quad\quad \text{By the definition of $\phi$}& \\
  &{\bf let \: pure \:} \vec{x} = \phi(\vec{M}[y := P]) {\: \bf in \:} \phi(N) = &\\
  &\quad\quad\quad\quad\quad\quad \text{Induction hypothesis}& \\
  &{\bf let \: pure \:} \vec{x} = (\phi(\vec{M}) [y := \phi(P)]) {\: \bf in \:} \phi(N) =&\\
  &\quad\quad\quad\quad\quad\quad \text{Substitution definition} & \\
  &({\bf let \: pure \:} \vec{x} = \phi(\vec{M}) {\: \bf in \:} \phi(N))[y := \phi(P)]
  \end{array}$

@@

Strong functor on a monoidal category is an endofunctor equipped with a natural transformation so called tensorial strength:

\begin{center}
$\begin{array}{lll}
  \tau_{A, B} : A \otimes \mathcal{K}B \rightarrow \mathcal{K}(A \otimes B)
\end{array}$
\end{center}

@@

\xymatrix{
&&& \mathds{1} \otimes \mathcal{K}A \ar[drr]_{R_{\mathcal{K}A}} \ar[rr]^{\mu_{\mathds{1}, A}} && \mathcal{K}(\mathds{1} \otimes A) \ar[d]^{\mathcal{K}(R_A)}\\
&&&&& \mathcal{K}A
}
\end{defin}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma, x : A \vdash M : B]\!] = [\![M]\!] : [\![\Gamma]\!] \times [\![A]\!] \rightarrow [\![B]\!]$}
\UnaryInfC{$[\![\Gamma \vdash (\lambda x. M) : A \to B]\!] = \Lambda([\![M]\!]) : [\![\Gamma]\!]
\rightarrow[\![B]\!]^{[\![A]\!]}$}
\end{prooftree}

@@

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A ]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A]\!]$}
\AxiomC{$[\![\Gamma \vdash N : B ]\!] = [\![N]\!] : [\![\Gamma]\!] \rightarrow [\![B]\!]$}
\BinaryInfC{$[\![\Gamma \vdash \langle M, N \rangle : A \times B]\!] = \langle [\![M]\!], [\![N]\!] \rangle : [\![\Gamma]\!] \rightarrow
[\![A]\!] \times [\![B]\!]$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$[\![\Gamma \vdash M : A_1 \times A_2]\!] = [\![M]\!] : [\![\Gamma]\!] \rightarrow [\![A_1]\!] \times
[\![A_2]\!]$}
\RightLabel{$i \in \{1,2\}$}
\UnaryInfC{$[\![\Gamma \vdash \pi_i M : A_i]\!] = [\![\Gamma]\!] \xrightarrow{[\![M]\!]} [\![A_1]\!] \times
[\![A_2]\!] \xrightarrow{\pi_i} [\![A_i]\!]$}
\end{prooftree}

@@

$\begin{array}{lll}
&[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline {\quad} {\: \bf in \:} M : {\bf K} A]\!] = & \\
&\quad\quad\quad\quad\quad\quad \text{By interpretation}& \\
&\mathcal{K}([\![M]\!]) \circ u_{\mathds{1}} = & \\
&\quad\quad\quad\quad\quad\quad \text{By the definition of an applicative functor}& \\
&\mathcal{K}([\![M]\!]) \circ \eta_{\mathds{1}} = & \\
&\quad\quad\quad\quad\quad\quad \text{By naturality of $\eta$} & \\
&\eta_{[\![A]\!]} \circ [\![M]\!] = &\\
&\quad\quad\quad\quad\quad\quad \text{By interpretation} & \\
&[\![\vdash {\bf pure \:} M : {\bf K} A]\!]&
\end{array}$
\end{proof}

@@

$\begin{array}{lll}
&\eta_{\hat{B}} \circ f = & \\
&\quad\quad\quad\quad\quad\quad \text{By the definition}&\\
&[y, {\bf pure \:} y] \circ [x, M] = &\\
&\quad\quad\quad\quad\quad\quad \text{By the definition of composition} & \\
&[x, {\bf pure \:} y [y := M]] = &\\
&\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
&[x, {\bf pure \:} M]& \\
& &\\
&\text{On the other hand:}& \\
&\text{fmap } f \circ \eta_{\hat{A}} = & \\
&\quad\quad\quad\quad\quad\quad \text{By the definiton}& \\
&[z, {\bf let \: pure \:} x = z {\: \bf in \:} M] \circ [x, {\bf pure \: x}] = &\\
&\quad\quad\quad\quad\quad\quad \text{By the definition of composition}& \\
&[x, {\bf let \: pure \:} x = z {\: \bf in \:} M [z := {\bf pure \:} x]] = & \\
&\quad\quad\quad\quad\quad\quad \text{By substitution}& \\
&[x, {\bf let \: pure \:} x = {\bf pure \:} x {\: \bf in \:} M] = & \\
&\quad\quad\quad\quad\quad\quad \text{$\beta$-reduction rule}& \\
&[x, {\bf pure \:} M [x := x]] = &\\
&\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
&[x, {\bf pure \:} M]&
\end{array}$

@@

$\begin{array}{lll}
&\ast_{\hat{A}, \hat{B}} \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
&\quad\quad\quad\quad\quad\quad \text{By definition}&\\
&[p^{'}, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle] \circ [p, \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle] = & \\
&\quad\quad\quad\quad\quad\quad \text{By composition} & \\
&[p, {\bf let \: pure \:} x, y = \pi_1 p^{'}, \pi_2 p^{'} {\: \bf in \:} \langle x, y \rangle [p^{'} := \langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle]] = & \\
&\quad\quad\quad\quad\quad\quad \text{By substitution} & \\
&[p, {\bf let \: pure \:} x, y = \pi_1 (\langle {\bf pure \:} (\pi_1 p), \pi_2 p \rangle), \pi_2 (\langle \pi_1 p, {\bf pure \:} (\pi_2 p) \rangle) {\: \bf in \:} \langle x, y \rangle] = & \\
&\quad\quad\quad\quad\quad\quad \text{By $\beta$-reduction rules} & \\
&[p, {\bf let \: pure \:} x, y = {\bf pure \:} (\pi_1 p), \pi_2 p {\: \bf in \:} \langle x, y \rangle]&
\end{array}$

@@

  \begin{itemize}
    \item $\text{fmap } \alpha_{\hat{A},\hat{B},\hat{C}} \circ \tau_{\hat{A} \times \hat{B}, \hat{C}} = \tau_{\hat{A}, \hat{B} \times \hat{C}} \circ (id_{\hat{A}} \times \tau_{\hat{B}, \hat{C}}) \circ \alpha_{\hat{A}, \hat{B}, {\bf K}\hat{C}}$;
    \item ${\bf K}(R_{\hat{A}}) \circ \tau_{\mathds{1}, \hat{A}} = R_{{\bf K}\hat{A}}$.
  \end{itemize}
  where $\alpha_{\hat{A},\hat{B},\hat{C}} = [p, \langle \pi_1 (\pi_1 p), \langle \pi_2 (\pi_1 p), \pi_2 p \rangle \rangle]$ and
  $R = \pi_2$.
\end{lemma}

@@

\section{Acknowledgment}

  We are grateful to Valeria de Paiva, Neel Krishnaswami and Vladimir Vasyukov for consulting and advice.

  Special thanks to Vladimir Krupski for general idea, critics and support.
